---
title: "Data Import"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```



This document will show how to import data.



## Import the FAS Litters CSV

Import the dataset and clean the names.

```{r}
litters_df = read_csv("data/FAS_litters.csv")

litters_df = janitor::clean_names(litters_df)
```

## Look at the dataset

```{r}
litters_df

head(litters_df)
tail(litters_df, 10)
```

```{r eval = FALSE}
view(litters_df)
```



#Import FAS Pups

Use relative paths.

```{r}
pups_df = read_csv("data/FAS_pups.csv")

pups_df = janitor::clean_names(pups_df)
```



Use absolute path. Caution: If you move the folder, the code below will not work. Use relative paths instead.

```{r, eval = FALSE}
pups_df = read_csv("/Users/mahdi/Desktop/data_wrangling_I/data/FAS_pups.csv")
```


## Look at read_csv options

col_names and skip rows

Skip = will be helpful if you don't want to include first couple of rows

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    col_names = FALSE,
    skip = 1
  )
```


What about missing data

using na = to correct missing variables

```{r}
litters_df =
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", ".")
  )

litters_df = janitor::clean_names(litters_df)
```


What if we code `group` as a factor variable?

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", "."),
    col_types = cols(
      Group = col_factor()
    )
  )
```


## Import an excel file

Import MLB 2011 summary data using read_excel, sheet, range.

```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet = "mlb11")
```


## Import SAS data

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```



## Never use read.csv()

Use `read_csv()`

```{r}
litters_df = read.csv("data/FAS_litters.csv")
```

Never do this either:
`$` are generally bad
```{r}
litters_df$L
```


