---
title: "Data Manipulation"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```


This document will show how to *manipulate* data.



Import the two datasets that we're going to manipulate.

```{r}
litters_df =
  read_csv("data/FAS_litters.csv", na = c("NA", "", "."))

litters_df = janitor::clean_names(litters_df)

pups_df =
  read_csv("data/FAS_pups.csv", na = c("NA", "", "."))

pups_df = janitor::clean_names(pups_df)
```


## `select`

Use `select()` to select variables (columns).

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

This will select everything from one thing:next

```{r}
select(litters_df, group:gd18_weight)
```

```{r}
select(litters_df, -pups_survive)
```

```{r}
select(litters_df, -(group:gd18_weight))
```

```{r}
select(litters_df, starts_with("gd"))
```

```{r}
select(litters_df, contains("pups"))
```

Use `rename()` to rename a variable but keep all the other variables.
Use `relocate()` to move variables around but keep everything else the same.

```{r}
select(litters_df, GROUP = group)

rename(litters_df, GROUP = group)

select(litters_df, litter_number, gd0_weight, everything())

relocate(litters_df, litter_number, gd0_weight)
```


Learning Assessment: in pups data, select litter number, sex, and pd_ears.

```{r}
select(pups_df, litter_number, sex, pd_ears)
```


## `filter`

Use `filter()` to filter out ROWS.
`==` means that is equal to (value that you want that is equal to this)

```{r}
filter(litters_df, gd_of_birth == 20)
filter(litters_df, gd_of_birth == 19)
```

```{r}
filter(litters_df, pups_born_alive > 8)
filter(litters_df, pups_born_alive >= 8)
```

```{r}
filter(litters_df, pups_born_alive != 9)
```

```{r}
filter(litters_df, group == "Con7")
filter(litters_df, group == "Low8")
filter(litters_df, group %in% c("Low7", "Low8"))
```

```{r}
filter(litters_df, group %in% c("Low7", "Low8"), pups_born_alive == 8)
```

Get rid of rows that have missing values using `drop_na()`

```{r}
drop_na(litters_df)

drop_na(litters_df, gd0_weight)

```

Learning Assessment: Filter to include only pups with sex 1 and Filter to include only pups with PD walk less than 11 and sex 2.

```{r}
filter(pups_df, sex == 1)

filter(pups_df, sex == 2, pd_walk < 11)
```


## `mutate`

You can create new variables by manipulating other variables.
You can change the characters data within variables.

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, sq_pups = pups_born_alive^2)

mutate(litters_df, group = str_to_lower(group))

mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```


## `arrange`

Can arrange data in ascending or descending order.

```{r}
arrange(litters_df, gd0_weight)

arrange(litters_df, desc(gd0_weight))

arrange(litters_df, pups_born_alive, gd0_weight)
```


## PIPING!!!!!

"and then" = command shift M *|>*
Pipe in logical order. 

```{r}
litters_df =
  read_csv("data/FAS_litters.csv", na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  select(-pups_born_alive) |>
  filter(group == "Con7") |> 
  mutate(
    wt_gain = 
      gd18_weight - gd0_weight,
    group = str_to_lower(group)
  )
```





